Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-01-28 doc/manuscript/ESSDmanuscript.Rmd
add a reference
paper edits

**As we seek to mitigate climate change [@unfccc_adoption_2015], forests are critical.**
*(See intros to other papers for relevant refs)*
Under Paris Agreement, **##**% of Nationally Determined Contributions relate to forests [@grassi_key_2017].

**The International Panel on Climate Change (IPCC) provides guidance for national greenhouse gas (GHG) inventories for reporting to the United Nations Framework Convention on Climate Change [UNFCCC, (REFS for older guidelines), @ipcc_2019_2019].**
Under this guidance, GHG inventories include all managed land, including most of the world's forest land [@ogle_delineating_2018].
The IPCC inventory guidelines include specific instructions for accounting for forest land ... [@ipcc_agriculture_2006; @ipcc_2019_2019a].
This guidance has improved as more of the relevant underlying data has become available [@requenasuarez_estimating_2019], but there remains room for continuous improvement as the science advances.
For example, @cook-patton_mapping_2020 *(summarize IPCC comparison results)*. 
*Moreover, it is useful for those compiling national greenhouse gas inventories to have access to locally-specific information, when available.*
To improve the data accessible for C accounting, the IPCC created the Emission Factor Database (EFDB; https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a recognized library of emission factors and other parameters that can be used for estimating greenhouse gas emissions and removals.

>>>>

Forests are critical to our management of greenhouse gases. The influence the forests have on asmosphereic carbon has made forest data increasingly valuable.

**The International Panel on Climate Change (IPCC) provides guidance for national greenhouse gas (GHG) inventories for reporting to the United Nations Framework Convention on Climate Change [UNFCCC, (REFS for older guidelines), @ipcc_2019_2019].**
Under this guidance, GHG inventories include all managed land, including most of the world's forest land [@ogle_delineating_2018].
The IPCC inventory guidelines include specific instructions for accounting for forest land ... [@ipcc_agriculture_2006; @ipcc_2019_2019a].
This guidance has improved as more of the relevant underlying data has become available [@requenasuarez_estimating_2019], but there remains room for continuous improvement as the science advances.
For example, @cook-patton_mapping_2020 *(summarize IPCC comparison results)*. 
Another example is @cuni-sanchez_high_2021.
*Moreover, it is useful for those compiling national greenhouse gas inventories to have access to locally-specific information, when available.*
To improve the data accessible for C accounting, the IPCC created the Emission Factor Database (EFDB; https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a recognized library of emission factors and other parameters that can be used for estimating greenhouse gas emissions and removals.

### 2022-01-28 doc/manuscript/ESSDmanuscript.tex
paper edits

%% Journal abbreviations (please use the same for discussion papers and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% ASTRA Proceedings (ap)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Bone and Joint Infection (jbji)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Polarforschung - Journal of the German Society for Polar Research (polf)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

@@

% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
%%\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
%

\begin{document}

@@

\textbf{As we seek to mitigate climate change
\citep{unfccc_adoption_2015}, forests are critical.} \emph{(See intros
to other papers for relevant refs)} Under Paris Agreement,
\textbf{\#\#}\% of Nationally Determined Contributions relate to forests
\citep{grassi_key_2017}.

\textbf{The International Panel on Climate Change (IPCC) provides
guidance for national greenhouse gas (GHG) inventories for reporting to
the United Nations Framework Convention on Climate Change \citep[UNFCCC,
(REFS for older guidelines),][]{ipcc_2019_2019}.} Under this guidance,
GHG inventories include all managed land, including most of the world's
forest land \citep{ogle_delineating_2018}. The IPCC inventory guidelines
include specific instructions for accounting for forest land \ldots{}
\citep{ipcc_agriculture_2006, ipcc_2019_2019a}. This guidance has
improved as more of the relevant underlying data has become available
\citep{requenasuarez_estimating_2019}, but there remains room for
continuous improvement as the science advances. For example,
\citet{cook-patton_mapping_2020} \emph{(summarize IPCC comparison
results)}. \emph{Moreover, it is useful for those compiling national
greenhouse gas inventories to have access to locally-specific
information, when available.} To improve the data accessible for C
accounting, the IPCC created the Emission Factor Database (EFDB;
https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a
recognized library of emission factors and other parameters that can be
used for estimating greenhouse gas emissions and removals.

@@

\caption{\label{tab:table_variables}\textbf{Variables with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{pool} & \textbf{definition} & \textbf{major sources of estimate variation} & \textbf{IPCC guidance}\\
\hline
aboveground biomass & all biomass of living vegetation, both woody and herbaceous, above the soil & allometry, min dbh & acceptable to exclude understory\\
\hline
belowground biomass & all biomass of live roots & allometry, min dbh, assumed ratio of belowground to aboveground biomass (IPCC table 4.4) & fine roots may be excluded when they cannot be distinguished empirically from soil organic matter or litter\\
\hline
dead wood & all non-living woody biomass not contained in the litter, either standing, lying on the ground, or in the soil & min dbh,  ... & default min dbh = 10cm, but may be chosen by country\\
\hline
litter & all non-living biomass with a size greater than the limit for soil organic matter  and less than the minimum diameter chosen for dead wood, lying dead, in various states of decomposition above or within the mineral or organic soil & min dbh for dead wood , .. & includes entire O horizon\\
\hline
soil organic matter & organic carbon in mineral soils to a specified depth & sampling depth & default sampling depth = 30cm, but may be chosen by country\\
\hline
\end{tabu}
\end{table}

@@

\textbf{ForC table} & \textbf{ForC field} & \textbf{EFDB field} & \textbf{Usage} & \textbf{Required}\\

@@

Define relationship among NEE, NEP, and delta.C., especially noting role
of harvest \citep{chapin_reconciling_2006}.

@@

\begin{itemize}
\item
  adopt common standards for variables like min diameter of deadwood,
  select soil sampling increments to include a cutoff at 30.
\item
  report 95\% CIs, SE, or STD and n
\item
  report C variables in article text, table, or SI table. EFDB cannot
  accept data digitized from figures
\item
  present calculations of all variables that would be useful to IPCC.
  EFDB requires that data in the database be presented in the original
  article, and as such cannot accept subsequent calculations. For
  example, if aboveground biomass and total biomass are presented, but
  root biomass is not presented, root biomass cannot be subsequently
  calculated and sent to EFDB.
\end{itemize}

>>>>

%% Journal abbreviations (please use the same for preprints and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Bone and Joint Infection (jbji)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Polarforschung - Journal of the German Society for Polar Research (polf)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Safety of Nuclear Waste Disposal (sand)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

@@

% The "Technical instructions for LaTex" by Copernicus require _not_ to insert any additional packages.
% 
% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%%\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
%
\begin{document}

@@

Currently the world is in a climate crisis and as we seek to find
answers climate data is increasingly important.Forests are critical to
our management of greenhouse gases. The influence the forests have on
asmosphereic carbon has made forest data increasingly valuable.

\textbf{The International Panel on Climate Change (IPCC) provides
guidance for national greenhouse gas (GHG) inventories for reporting to
the United Nations Framework Convention on Climate Change \citep[UNFCCC,
(REFS for older guidelines),][]{ipcc_2019_2019}.} Under this guidance,
GHG inventories include all managed land, including most of the world's
forest land \citep{ogle_delineating_2018}. The IPCC inventory guidelines
include specific instructions for accounting for forest land \ldots{}
\citep{ipcc_agriculture_2006, ipcc_2019_2019a}. This guidance has
improved as more of the relevant underlying data has become available
\citep{requenasuarez_estimating_2019}, but there remains room for
continuous improvement as the science advances. For example,
\citet{cook-patton_mapping_2020} \emph{(summarize IPCC comparison
results)}. Another example is \citet{cuni-sanchez_high_2021}.
\emph{Moreover, it is useful for those compiling national greenhouse gas
inventories to have access to locally-specific information, when
available.} To improve the data accessible for C accounting, the IPCC
created the Emission Factor Database (EFDB;
https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a
recognized library of emission factors and other parameters that can be
used for estimating greenhouse gas emissions and removals.

@@

\caption{\label{tab:table_variables}\textbf{Variables with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{ï  pool} & \textbf{definition} & \textbf{major sources of estimate variation} & \textbf{IPCC guidance}\\
\hline
aboveground biomass & all biomass of living vegetation, both woody and herbaceous, above the soil & allometry, min dbh & acceptable to exclude understory\\
\hline
belowground biomass & all biomass of live roots & allometry, min dbh, assumed ratio of belowground to aboveground biomass (IPCC table 4.4) & fine roots may be excluded when they cannot be distinguished empirically from soil organic matter or litter\\
\hline
dead wood & all non-living woody biomass not contained in the litter, either standing, lying on the ground, or in the soil & min dbh,  ... & default min dbh = 10cm, but may be chosen by country\\
\hline
litter & all non-living biomass with a size greater than the limit for soil organic matter  and less than the minimum diameter chosen for dead wood, lying dead, in various states of decomposition above or within the mineral or organic soil & min dbh for dead wood , .. & includes entire O horizon\\
\hline
soil organic matter & organic carbon in mineral soils to a specified depth & sampling depth & default sampling depth = 30cm, but may be chosen by country\\
\hline
\end{tabu}
\end{table}

@@

\textbf{ï  ForC table} & \textbf{ForC field} & \textbf{EFDB field} & \textbf{Usage} & \textbf{Required}\\

@@

@@

\begin{itemize}
\item
  adopt common standards for variables like min diameter of deadwood,
  select soil sampling increments to include a cutoff at 30.
\item
  report 95\% CIs, SE, or STD and n
\item
  report C variables in article text, table, or SI table. EFDB cannot
  accept data digitized from figures
\item
  present calculations of all variables that would be useful to IPCC.
  EFDB requires that data in the database be presented in the original
  article, and as such cannot accept subsequent calculations. For
  example, if aboveground biomass and total biomass are presented, but
  root biomass is not presented, root biomass cannot be subsequently
  calculated and sent to EFDB. Similarly, fine and coarse root biomass
  can't be summed; soil carbon can't be summed across depth increments,
  etc.
\end{itemize}
